### 2011-12-31 trunk/GUIErrorDetector/paper/abstract.tex
tweak

In this paper, we present a general technique to find such \textit{invalid thread access}
errors in multithreaded GUI applications. We formulate finding invalid
thread access errors as a call graph
reachability problem. Our technique starts from call graph entry nodes and statically explores
all possible paths that may cause a non-UI thread to access a GUI object.
However, the heavy use of reflection in modern GUI frameworks makes existing
call graph construction algorithms insufficent to build a both complete and precise
call graph. To overcome such limitations, we present a combined RTA and k-CFA
call graph analysis (that XXX). % to overcome such limitations.% of analyzing GUI applications.

We implemented our technique and instantiated it on four popular GUI
platforms, namely SWT, Eclipse plugin development framework, Java Swing, and Android. In an evaluation
on XX subjects over XXX LOC, our technique found a number of errors,
suggesting its effectiveness and showing good scalability.

>>>>

In this paper, we present a general technique to find such \textit{invalid thread access}
errors in multithreaded GUI applications. We formulate finding invalid
thread access errors as a call graph
reachability problem. Our technique starts from call graph entry nodes to
statically explore all possible paths that may cause a non-UI thread
to access a GUI object.
However, the heavy use of reflection in modern GUI applications makes existing
call graph construction algorithms insufficient to build a both complete and precise
call graph. Thus, we present a combined RTA and k-CFA
call graph construction algorithm to overcome the limitations.% of analyzing GUI applications.

We implemented our technique and instantiated it on four popular GUI
platforms, namely SWT, Eclipse plugin development framework, Java Swing, and Android. In an evaluation
on XX subjects over XXX LOC, our technique found a number of real-world errors,
suggesting its effectiveness and showing good scalability.

### 2011-12-31 trunk/GUIErrorDetector/paper/technique.tex
tweak

At a high level, to detect an invalid thread access error, an analysis only needs to track all
non-UI threads spawned by the UI thread, and check whether those non-UI threads
may invoke a UI accessing method.

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specially, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretical ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative static call graph is a superset of the ideal call graph;
it over-approximates the
dynamic call graph of every possible execution. 
We believe call graph has a suitable granularity for this problem while
Compared with other
program representations like control flow graph or program dependence
graph. That is because, in an object-oriented programming language like
Java, a method is the basic unit of a program. Thus, reporting errors
at the method level can be sufficient in most cases.

Our algorithm in Figure~\ref{fig:detectalgorithm} works as follows. It
constructs a static call graph for the tested program (line 2),
and specifies entry nodes for it. The algorithm starts to perform
call graph traverse from every entry node (line 6). For every entry node,
it first uses Breadth First Search (BFS) to find all reachable
\CodeIn{Thread.start()} method (line 7). Each \CodeIn{Thread.start()} method
indicates a new, non-UI thread is spawned.  Then, from each
\CodeIn{Thread.start()} node, the algorithm uses a slightly modified 
BFS algorithm to find reachable UI accessing methods (lines 9 -- 23).
Specifially, it uses a queue to keep track of all visited nodes, and
the algorithm checks every successive node to be visited;
it reports an error if that node
is a UI accessing method (line 15), or skips it if it is
a safe method (line 17), or simply keeps traversing (line 19).

@@

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. Existing
call graph construction algorithms either suffer from incompleteness
or imprecision. GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points, therefore, reflection must be properly handled.
We developed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}
to alleviate this problem.

We also note that static analysis may report false positives due
to its conservative nature, and multiple warnings that acutally corrspond to
the same error. In addition, some GUI applications may even use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
must be considered to make the call graph sound. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and annotation
support for programmers to specify call relationship for native methods.

@@

In the context of object-oriened languages wth dynamic dispatch, the
crucial step in constructing a call graph is to compute a conservative
approximation of the set of methods that can be invoked by a given
virtual method call. Call graph construction algorithms have been
studied intensively in the 1990s. While their original formulations
use a variety of formalisms, most of them can be recast as set-based
analyses. Various algorithms differ primarily in the number of sets
that are used to approximate run-time values of expressions. 
In practice, algorithms such as RTA that use a single set for the whole
program scale well but achieves relatively lower precision. The
k-CFA algorithms maintains one set per expression achieves better
precision than RTA.

Unfortunately, for GUI applications that use reflection, we found both
RTA and k-CFA is insufficient to build a satisfactory call graph. Here
is an example in Figure~\ref{fig:sampleandroid} to illustrate their limitations.

(XXX)

For desktop applications which have main method, it is straigtforward to apply
the above algorithm. However, for applications on platforms like eclipse plugin
and Android. presents new challenges. Two technical challenges must be solve.
First, use reflection extensively, Second, it is possible to interact with
native code.

Android features an extended event-based library and dynamic inflation of
graphical views from declarative XML layout files.
A static analyzer for Android programs must consider such features, for
correctness and precision.

Give an example of how reflection affect call graph precision.

The Android button example.

Using RTA is not enough, since that maintains a global

So, we present a new call graph construction algorithm

reflected classes are known, use RTA.

reflection code

native code

\begin{figure}[t]
\textbf{Input}: a Java program $P$, a list of classes $cls$\\
\textbf{Output}: a call graph $cg$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE callSiteMap $\leftarrow$ new Map$\langle$CallSite, List$\langle$Type$\rangle$$\rangle$
\STATE rta\_cg $\leftarrow$ $constructCallGraphByRTA$($P$)
\FOR{each $method$ in $P$}
\FOR{each  $callSite$ in $method$}
\STATE type $\leftarrow$ $staticType$(callSite)
\IF{type $\in$ $cls$}
\FOR{each subType of $subType$(type)}
\IF{$hasObjectCreated$(subType, rta\_cg)}
\STATE callSiteMap[callSite].add(subType)
\ENDIF
\ENDFOR
\ENDIF
\ENDFOR
\ENDFOR
\STATE $cg$ $\leftarrow$ $constructCallGraphByKCFA$($P$, callSiteMap)
\RETURN $cg$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{A combined  RTA~\cite{rta} and k-CFA~\cite{kcfa} call graph construction algorithm.
$cls$ is a list of classes whose method invocations are added to the
call graph using the RTA algorithm; while method invocations from other classes
are added to the call graph using the k-CFA algorithm.
} \label{fig:cgalgorithm}
\end{figure}

>>>>

At a high level, to detect an invalid thread access error, an analysis needs to track all
non-UI threads spawned by the UI thread, and check whether those non-UI threads
may invoke a UI accessing method.

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specially, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretical ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative static call graph used in practice is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
In an object-oriented programming language like
Java, a method is the basic logic unit in a program. Reporting
potential errors at the method level can be sufficient in most cases.
Thus, we believe using call graph is well suited for this problem while
compared with other
program representations like control flow graph or program dependence
graph.

Our algorithm in Figure~\ref{fig:detectalgorithm} works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes for it (line 3). The algorithm starts to traverse
the call graph from every entry node (line 6). For every entry node,
it first uses Breadth First Search (BFS) to find all reachable
\CodeIn{Thread.start()} method (line 7). Each \CodeIn{Thread.start()} method
indicates that a new, non-UI thread is spawned.  After that, 
the algorithm uses a slightly modified 
BFS algorithm to find reachable UI accessing methods (lines 9 -- 23)
from each \CodeIn{Thread.start()} node.
Specifially, it uses a queue to keep track of all visited nodes, and
then checks every successive node to be visited (line 12);
it reports an error if that node
is a UI accessing method (line 15), or skips it if it is
a safe method (line 17), or simply keeps traversing otherwise (line 19).

@@

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. 
In particular, GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points.
Existing call graph construction algorithms either suffer from incompleteness
or imprecision.  To alleviate this problem, 
We proposed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}.

We also note that a static analysis may report false positives due
to its conservative nature, or multiple warnings that acutally corrspond to
the same error. In addition, some GUI applications may use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
should be considered to construct a good call graph. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and provide
annotation support for programmers to specify call relationship for
 native methods.

@@

In the context of object-oriened languages wth dynamic dispatch, the
crucial step in constructing a call graph is to compute a conservative
approximation of the set of methods that can be invoked by a given
virtual method call. Call graph construction algorithms have been
studied intensively in the 1990s. While their original formulations
use a variety of formalisms, most of them can be recast as set-based
analyses. Various algorithms differ primarily in the number of sets
that are used to approximate run-time values of expressions. 
In practice, algorithms such as RTA that use a single set for the whole
program scale well but achieves relatively lower precision. The
k-CFA algorithms maintains one set per expression achieves better
precision than RTA (\textit{k} is the context length, which is
orthogonal to the number of maintained sets).

\begin{figure}[t]
\textbf{Input}: a Java program $P$, a list of classes $cls$\\
\textbf{Output}: a call graph $cg$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE callSiteMap $\leftarrow$ new Map$\langle$CallSite, List$\langle$Type$\rangle$$\rangle$
\STATE rta\_cg $\leftarrow$ $constructCallGraphByRTA$($P$)
\FOR{each $method$ in $P$}
\FOR{each  $callSite$ in $method$}
\STATE type $\leftarrow$ $staticType$(callSite)
\IF{type $\in$ $cls$}
\FOR{each subType of $subType$(type)}
\IF{$hasObjectCreated$(subType, rta\_cg)}
\STATE callSiteMap[callSite].add(subType)
\ENDIF
\ENDFOR
\ENDIF
\ENDFOR
\ENDFOR
\STATE $cg$ $\leftarrow$ $constructCallGraphByKCFA$($P$, callSiteMap)
\RETURN $cg$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\label{fig:cgalgorithm}
\caption{A combined  RTA~\cite{rta} and k-CFA~\cite{kcfa} call graph construction algorithm.
$cls$ is a list of classes whose method invocations are added to the
call graph using the RTA algorithm; while method invocations from other classes
are added to the call graph using the k-CFA algorithm.
} 
\end{figure}

Unfortunately, GUI applications often use reflection, which makes either
RTA or k-CFA insufficient to build a satisfactory call graph. 
The example code in Figure~\ref{fig:sampleandroid}  illustrates their limitations.
In Figure~\ref{fig:sampleandroid}, line 7 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the XML file. Due to
its limitation in handling reflection, the k-CFA algorithm
will conclude that variable \CodeIn{button} at line 7 does not point to any
concrete \CodeIn{Button} object. Thus, it will miss the call edge \CodeIn{setText}
while processing line 11. On the other hand, a less precise algorithm like RTA~\cite{}
can not solve this problem. By explicitly specifying that GUI objects
like \CodeIn{Button} as declared in the XML file will be created, RTA maintains
a global value set for those GUI object type, and concludes that the \CodeIn{button} variable
at line 7 points to a concrete Button object. So the algorithm  adds a
call edge \CodeIn{setText} to the call graph. However, using RTA for
the whole program results in unnecessary precision loss. Note that,
RTA may conclude that the \CodeIn{cachedViews} variable at line 2
may point to any created \CodeIn{List} objects, despite the
fact that \CodeIn{cachedViews} can only be a \CodeIn{LinkedList} object at line 12.

To alleviate the above problems, we present a combined RTA and k-CFA
call graph construction algorithm as shown in Figure~\ref{fig:cgalgorithm}.
This algorithm uses the RTA algorithm to construct call graph for part of the program that
uses reflection to creating GUI objects (to ensure call graph completeness), while still uses
the k-CFA algorithm for the remaining part of the program (to avoid precision loss). Specifically,
users provide a list of classes, which can either be encoded manually or
automatically extracted from the XML configuration file, whose method invocations should be
constructed in a RTA way. The algorithm maintains a global set for
each  provided class, but a separate set for expressions of other
types.  For the example in Figure~\ref{fig:sampleandroid}, this
combined algorithm concludes that a \CodeIn{Button} object has been
created at line 7 and then adds the call edge \CodeIn{setText} to the
graph. It also concludes that the \CodeIn{cachedViews} variable can only
point to a \CodeIn{LinkedList} object, avoiding unnecessary precision loss.

%For desktop applications which have main method, it is straigtforward to apply
%the above algorithm. However, for applications on platforms like eclipse plugin
%and Android. presents new challenges. Two technical challenges must be solve.
%First, use reflection extensively, Second, it is possible to interact with
%native code.

%Android features an extended event-based library and dynamic inflation of
%graphical views from declarative XML layout files.
%A static analyzer for Android programs must consider such features, for
%correctness and precision.

%Give an example of how reflection affect call graph precision.

%So, we present a new call graph construction algorithm

%reflected classes are known, use RTA.

%reflection code

%native code
